sdmTMB: an R package for fast, flexible, and user-friendly generalized linear mixed effects models with spatial and spatiotemporal random fields

### 2023-02-14 doc/paper-jss/sdmTMB-cover.tex
Update cover

\documentclass[letterpaper]{letter}
\usepackage[margin=1.50in]{geometry}
\geometry{letterpaper}
\usepackage{graphicx}
\addtolength{\topmargin}{-30pt}
\addtolength{\textheight}{55pt}

@@

Thank you for considering our manuscript for publication in the Journal of Statistical Software. Our manuscript introduces the \proglang{R} package \pkg{sdmTMB} for fitting a broad class of generalized linear mixed effect models (GLMMs) that include spatial or spatiotemporal random fields. Such models are useful across a wide array of disciplines.

@@

Accessibility: \pkg{sdmTMB} provides a much more usable interface than alternatives, particularly for users who may be familiar with the syntax found in \proglang{R}'s \texttt{glm()}, \pkg{glmmTMB}, \pkg{lme4}, or \pkg{mgcv} packages. Existing packages, such as \pkg{INLA} or \pkg{inlabru} are fundamentally Bayesian and require users to deal with the complexity of priors. While \pkg{sdmTMB} can incorporate priors and extend to full Bayesian inference via \pkg{Stan}, by default estimation is done using the simplicity and speed of maximum likelihood.

Speed: The package combines the SPDE approximation with estimation in the \pkg{TMB} \proglang{R} package for computational efficiency. Across our comparisons, \pkg{sdmTMB} was 6--24 times faster than the \pkg{INLA} or \pkg{inlabru} packages, the software with the most similar capabilities for this class of models.

@@

JSS would be the ideal outlet to provide a citable resource for \pkg{sdmTMB} and to introduce the package to a wider audience. \pkg{sdmTMB} is an open-source \proglang{R} package \href{https://CRAN.R-project.org/package=sdmTMB}{published on CRAN}. Our code is fully reproducible and included with our submission. We have posted a preprint version of our manuscript \href{https://doi.org/10.1101/2022.03.24.485545}{on bioRxiv}. We look forward to constructive feedback from your reviewers and editors.

@@

>>>>

\documentclass[letterpaper]{letter}
\usepackage[margin=1.55in]{geometry}
\geometry{letterpaper}
\usepackage{graphicx}
\addtolength{\topmargin}{-30pt}
\addtolength{\textheight}{55pt}

@@

Thank you for considering our manuscript for publication in the Journal of Statistical Software. This is a revised version of manuscript 4939 after receiving the following editor comments with an encouragement to resubmit:

\begin{quote}
\input{editor-comments0}
\end{quote}

We apologize for not better dealing with the GitHub-only package \pkg{sdmTMBextra}. Indeed, this was originally a part of the main package. We had just split this functionality out of the main package because of valgrind/gcc-ASAN check errors on CRAN that could not be replicated or resolved. This also had the benefit of substantially reducing the number of package dependencies for the main CRAN package. We have now omitted the corresponding code in the appendix and referenced the vignette on residuals for further details.

Our original cover letter follows:

Our manuscript introduces the \proglang{R} package \pkg{sdmTMB} for fitting a broad class of generalized linear mixed effect models (GLMMs) that include spatial or spatiotemporal random fields. Such models are useful across a wide array of disciplines.

@@

Accessibility: \pkg{sdmTMB} provides a more usable interface than alternatives, particularly for users who may be familiar with the syntax found in \proglang{R}'s \texttt{glm()}, \pkg{glmmTMB}, \pkg{lme4}, or \pkg{mgcv} packages. Furthermore, existing packages such as \pkg{INLA} or \pkg{inlabru} are fundamentally Bayesian and require users to deal with the complexity of priors. While \pkg{sdmTMB} can incorporate priors and extend to full Bayesian inference via \pkg{Stan}, by default estimation is done using the simplicity and speed of maximum likelihood.

Speed: The package combines the SPDE approximation with estimation in the \pkg{TMB} \proglang{R} package for computational efficiency. Across our comparisons, \pkg{sdmTMB} was 6--24 times faster than the \pkg{INLA} or \pkg{inlabru} packages---the software with the most similar capabilities for this class of models.

@@

JSS would be the ideal outlet to provide a citable resource for \pkg{sdmTMB} and to introduce the package to a wider audience. \pkg{sdmTMB} is an open-source \proglang{R} package \href{https://CRAN.R-project.org/package=sdmTMB}{published on CRAN}. Our code is fully reproducible and included with our submission. We have posted a preprint version of our manuscript \href{https://doi.org/10.1101/2022.03.24.485545}{on bioRxiv}.  We have included a list of potential reviewers on the following page should it be helpful. We look forward to constructive feedback from your reviewers and editors.

@@

\vspace{6mm}

Suggested reviewers:

\input{suggested-reviewers}
